Breakable Elastic-plastic Constraint for\\Rigid Body Simulation

### 2017-08-04 pdocs/thesis/epj-jcgt.tex
work on jcgt article

%  Simo.Nikula@gmail.com
\def\mytitle{Breakable Elastic-plastic Constraint for Rigid Body Simulation}
\def\lut{Lappeenranta University of Technology}

@@

In impulse-based phycis engines most breakable scenarios  are implemented by destructing various bodies based on collision
or impulse exceeding predefined limit.
Nevertheless, breaking of steel or reinforced concrete structures using this approach 
is not appropriate if the simulation is to look realistic.

@@

\paragraph{Our approach}
This study will introduce an approach to account for plastic deformation in impulse-based phycis engines.   
Presented methodology does not require significant software development efforts from
game vendors and is thus easily adoptable. 
In the introduced method, plastic deformation takes place if the force or moment exceeds a predefined 
limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
Maximum forces and moments are estimated based on the plastic section modulus.
Joint breaking is based on summing plastic deformation and comparing it to a
predefined material based limit. The elastic part of deformation is modelled by employing 
modification of an existing constraints.

@@

\paragraph{Joint breaking}

\section{Implementation for Box2D}

@@

>>>>

%  Simo.Nikula@gmail.com
\def\mytitle{Breakable Elastic-plastic Constraint for Rigid Body Simulation with Impulse based Physics}
\def\lut{Lappeenranta University of Technology}

@@

In impulse-based phycis engines most breakable scenarios  are implemented by destructing various bodies based on collision
or impulse exceeding predefined limit. This provides simple means for visualization of explosions and high speed collisions 
on brittle structures.  
Nevertheless, breaking of steel or reinforced concrete structures using this approach 
is not appropriate if the simulation is to look realistic or structures should not fail completely.

@@

\paragraph{Our approach}
This study will introduce an approach to account for plastic deformation in impulse-based phycis engines.   
Presented methodology does not require significant software development efforts from
game vendors and is thus easily adoptable. 
In the introduced method, plastic deformation takes place if the force or moment exceeds a predefined 
limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
Maximum forces and moments can be estimated based on the plastic section modulus or
by defining maximum elastic displacement.
Joint breaking is based on summing plastic deformation and comparing it to a
predefined material and geometry based limit. The elastic part of deformation is modelled by employing 
modification of an existing constraints.

@@

\paragraph{Joint breaking}
Joint breaking is based on summing of plastic displacement or rotation.
Process can be subdivided few phases.

\begin{enumerate}
\item Hitting maximum impulse value triggers additional processing.
\item New neutral position is calculated so that elastic part is subtracted from current position.
\item Plastic strain or rotation is incremented.
\item Current plastic strain or rotation is compared to maximum value and joint is broken if 
current value exceeds predefined maximum value. This part can be finetuned e.g. so that joint is 
not broken during compression if connected parts are made of material which can handle large
strains due to compressive forces. 
\end{enumerate}

\section{Implementation for Box2D}
b2ElasticPlastic joint implementation is based on b2WeldJoint which implements rigid and flexible joints as
described in Catto's paper~\shortcite{ecsc} on soft constraints. Main idea is that game designer should not have
find suitable values to get desired functionality but constraint is configured by giving desired  damping ratio and frequency.
This principle leads in many cases to quite soft constraints to keep simulation stable. 
b2WeldJoint supports also rigid joint in which case iterative solver creates some flexibility.

@@

\begin{enumerate}
\item Box2D source with plasticity extensions 
 \href{https://github.com/simo-11/Box2D}{https://github.com/simo-11/Box2D}

\item \cbullet source with plasticity extensions 
 \href{https://github.com/simo-11/bullet3}{https://github.com/simo-11/bullet3}
\item two minute video demonstrating \cbullet implementation 
 \href{https://youtu.be/_QNbv1-f_ao}{https://youtu.be/_QNbv1-f_ao}
\item 2 MB executable used for \cbullet video {https://github.com/simo-11/bullet3/blob/master/pbin/App_PlasticityExampleBrowser_vs2013.exe?raw=true}{https://github.com/simo-11/bullet3/blob/master/pbin/App_PlasticityExampleBrowser_vs2013.exe?raw=true}
\end{enumerate}
